MSstatsPTM: Statistical relative quantification of post-translational modifications in bottom-up mass spectrometry-based proteomics

### 2022-08-16 main/MSstatsPTM_Main.tex
evaluation comments

\todo{I am wondering if 'basic' is the right vocabulary here. Would you like to switch to 'unadjusted ANOVA', 'adjusted ANOVA', 'unadjusted Limma' and 'adjusted Limma', everywhere in the text and in the figures? }

$MSstatsPTM$ before adjusting for changes in the unmodified protein corresponds to base $MSstats$ or $MSstatsTMT$, modeled on the peptide level \todo{as described in Suppl Table XXX, as appropriate for the experimental design}. $MSstatsPTM$ with the adjustment, \todo{described in Sec XXX}, was used without imputing missing values, and without Empirical Bayes moderation.

$Basic$ $ANOVA$, i.e. $ANOVA$ before adjusting for changes in the unmodified protein, was as in \eqref{ttest_model}. $ANOVA$ with the adjustment was modeled as in \eqref{ttest_ratio_model}. 
Finally, $basic$ $Limma$ (i.e. without $Limma$ adjustment) used the same model formula in \eqref{ttest_model}, while including a moderated variance estimation. $Limma$ with adjustment was modeled as in \eqref{ttest_ratio_model} including moderated variance estimation.  
All the evaluations were done at the FDR-adjusted p-value cutoff of $q=.05$. More details are in {\bf Supplementary Sec. \ref{supp-sec:experiments}}.

@@

\figref{sim1_fdr} summarizes the eFDR reported on Computer Simulation 1 dataset by $MSstatsPTM$, $ANOVA$, $Limma$ and the basic $MSstats$, $ANOVA$ and $Limma$ methods. The  simulation mimicked a ``clean" label-free group comparison experiment, not compromised by issues such as deviations from model assumptions, missing values and outliers. All the analyses were performed to control the eFDR at at most 5\%.  Yet, even under these favorable circumstances, the models that did not adjust for confounding from changes in overall protein abundance produced an excessive number of false positives. The versions of the models that accounted for the confounding produced error rates that were much better calibrated at the desired level. \todo{Since the simulation mimicked the assumptions of model XXX?, the error rates of this model were calibrated best.}

@@

\figref{sim2_acc} summarizes the overall accuracy of methods adjusting for changes in the unmodified protein for Computer Simulation 2. The simulation mimicked a more realistic label-free group comparison experiment, including low counts of modified features and missing values. We evaluated the impact of low versus high noise, of the number of biological replicates, and of the number of conditions. For low noise, $MSstatsPTM$ outperformed the existing methods across all conditions and number of replicates, with near 100\% accuracy when the replicates were high. As the noise increased the accuracy of all the methods decreased, however $MStatsPTM$ still outperformed the existing methods. The difference is primarily due to the robust nature of summarization in $MSstatsPTM$ through Tukey's Median Polish. \todo{In Background we also said that a disadvantage of the ratios is in losing data when a ratio could not be constructed. Could you comment on how much less data was used by ratio-based approaches, as compared to our approach?}

@@

\figref{spikein_boxplot} summarizes the evaluation on the label-free spike-in group comparison experiment, for all the methods, with and without adjusting for changes in the unmodified protein. 
Without adjusting for changes in unmodified protein abundance, all the approaches incorrectly estimated the log$_2$-fold change of the modified spike-in peptides.
After adjustment, the estimation was generally in line with the ground truth for all methods, however $MSstatsPTM$'s distribution of estimated fold changes was tighter. 
\todo{On average over what?} the inter-quartile range of the estimates by $MSstatsPTM$ was 32.86\% smaller than that by $Limma$ or $ANOVA$. 
As seen in the previous section, the robust summarization by $MSstatsPTM$ was particularly useful when the number of \todo{features? Biological replicates?} was low. 
This was also the case in this experiment, with an average of $1.37$ features per PTM.

\figref{spikein_prop_volcano} details the detection of differentially abundant PTM for $MSstatsPTM$ and $Limma$, with and without the adjustment for changes in the unmodified protein, \todo{for which comparisons?}.
As above, the log$_2$-fold changes of the spike-in peptides were only correctly estimated when accounting for changes in the unmodified protein abundance. 
Additionally, the background peptides, serving as the null model, show many false positives before adjustment. 
After adjustment the extent of false positives substantially decreased. 
Specifically, for $MSstatsPTM$ the number of false positives went from 20.88\% to 1.84\% after the adjustment. \todo{the numbers for Limma?} 
While the proposed method and $Limma$ both correctly estimated the fold change of the spike-in peptides, using $Limma$ resulted in many large adjusted p-values, and lower sensitivity. 
This experiment introduced more variance components \todo{not sure what you mean. The design is group comparison, and the model has only one variance component. Rephrase?}. 
On average \todo{over all the proteins in this experiment?}, the variance components estimates by $Limma$ were 35.7\% larger than for $MSstatsPTM$. Volcano plots for all methods and comparisons can be seen in {\bf Supplementary Section~\ref{supp-sec:benchmark}}.

@@

\todo{Could you separate the stories? One paragraph for one dataset; another paragraph (or two?) for the other dataset.}The two results of these experiments are summarized in~\figref{venn_diagrams} and~\figref{profile_plots}. \figref{venn_diagrams} shows the number of differentially abundant modified peptides before and after adjustment for Datasets 4 \& 5. Adjusting for changes in the unmodified protein caused more PTMs to change from differentially abundant to non-differentially abundant than vice versa. A question that must be addressed is if the decrease in differentially abundant peptides is due to the increased variance \todo{variance of what? estimation of the variance parameter? Clarify} that comes from adjustment. This was tested by looking for modified peptides whose adjusted log fold change was within 10\% of the unadjusted log fold change but became statistically insignificant after adjustment. In other words, the fold change was the same between models but variance increased. When the test was applied to Dataset 4, only one PTM became non-differentially abundant due to an increase in variance. When applied to Dataset 5, 548 PTMs became non-differentially abundant due to an increase in variance (only 3.4\% of all PTMs that became non-differentially abundant). Thus we can conclude that the drop off in differentially abundant PTMs was due to changes in global protein abundance, not simply increased model variance.

\figref{venn_diagrams} demonstrates that the unmodified protein was more likely to cause the observed changes in PTM abundance, rather than mask changes in the PTM. The later case can occur when the unmodified protein exhibits a change in abundance, while the modification does not. Additionally, this case may come about due to the high variability of PTMs compared to the global protein. An example of this can be seen in \figref{data5_profile_plot}. Originally the modification at $S178$ of protein $TTP$ was shown to be differentially abundant between KO\_Uninfect and KO\_Early, with a log fold change of $2.9$. However the unmodified protein was shown to contribute $2.014$ of this change, meaning that the modification only accounted for $.886$ after adjustment (seen in {\bf Supplementary Figure~\ref{supp-fig:No_Diff_Shigella_PTM}}). This caused the modification to go from differentially abundant to non-differentially abundant. These cases occur frequently and can be thought of as false positives. The proposed approach corrects for this confounding, allowing us to see the true fold change of PTMs, while excluding those that do not exhibit differential changes.

Conversely, cases where the unmodified protein masks changes in the PTM are more infrequent. In this case, the modification and unmodified protein exhibit contradictory changes in abundance, moving in opposite directions. One such case can be seen in~\figref{data4_profile_plot}. Luchetti et al.~\cite{LUCHETTI2021} showed that $GSDMD$ was actively degraded when IpaH7.8 expression was induced by dox treatment. Our reanalysis confirmed this, with the $GSDMD$ protein being down-regulated when Dox treatments reached the 4 and 6 hour marks. Conversely, ubiquitination of $GSDMD$ at site $K62$ up-regulated abundance between the same conditions. This up-regulation was originally confounded by the down-regulation of unmodified $GSDMD$, and made the modification appear to have little change between no Dox and Dox 4 and 6 hour conditions. The proposed approach accounted for this confounding and the modification was shown to be differentially abundant, with an absolute fold change of $1.415$ between the Dox 1 hour and Dox 4 hour conditions. This was an interesting result, as the modification contradicts what previous research showed for the unmodified protein. This effect can be thought of as a false negative and would have been challenging to observe without the proposed approach. There are many non-differentially abundant PTMs, and having to comb through each one individually to determine false negatives is generally an unfeasible task. The proposed approach removes the confounding on an experimental level, revealing modifications of interest that are entirely masked due to changes in the unmodified protein.

@@

The number of differentially abundant PTMs before and after adjustment are shown in~\figref{data6_plots}. After adjusting for changes in the unmodified protein, there were fewer differentially modified peptides. However, this was mainly due to not having a global profiling run. When looking only at PTMs with a corresponding unmodified protein, the number of differentially modified PTMs exhibited a small drop, but was more consistent. As above, we tested if this drop in differentially abundant PTMs was due to an increase in variance. Here only 25 PTMs became insignificant due to increased variance. Without a global profiling run, the method was less effective at removing confounding with the unmodified protein. This was due to not being able to perform adjustment without a corresponding global protein. In this case adjustment caused us to toss out a large number of the PTMs in the original model. In cases without a global profiling run, the proposed method can still be used, but it should be used in conjunction with the unadjusted model so that information is not lost.

@@

\begin{figure}[ht]
\centering
\begin{subfigure}[c]{0.825\linewidth}
\includegraphics[width=1\textwidth]{images/spike_in_fc.png}
\caption{}
\label{fig:spikein_boxplot}
\end{subfigure}
\begin{subfigure}[c]{0.825\linewidth}
\includegraphics[width=1\textwidth]{images/spike_in_volcano.png}
\caption{}
\label{fig:spikein_prop_volcano}
\end{subfigure}
\caption{Dataset 3: Spike-in benchmark - Ubiquitination - Label-free. \todo{Describe the details of what we see. The high-level conclusions are best moved to the main text} a) \todo{First explain what we see in the figure, before interpreting the conclusions} Before adjustment the fold change of the spike-in peptides' were systematically different from the expected fold change in all models. After adjustment, this systemic difference was removed, however the inner quartile range of the $Limma$ and $ANOVA$ models was wider than the proposed method. b) \todo{explain the meaning of vertical solid vs dashed lines; and of the horizontal line; explain that only a subset of comparisons and methods}  Before adjustment the spike-in peptides (colored red) did not follow the expected log fold change; after adjustment, the spike-in peptides were more in line with expectation. Using $Limma$, the spike-in peptides followed the expected log fold change after adjustment, however the adjusted p-value for the majority of spike-in peptides were below the cutoff.}
\label{fig:spikein_volcano}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{\textwidth}
 \centering
\includegraphics[width=1.0\textwidth]{images/No_Difference_Shigella_Profile_Plot}
\caption{}
\label{fig:data4_profile_plot}
 \end{subfigure}
 \begin{subfigure}{\textwidth}
 \centering
	\includegraphics[width=1.0\textwidth]{images/IpaH_prof_plot.png}
	\caption{}
	\label{fig:data5_profile_plot}
	 \end{subfigure}
\caption{ \todo{Describe the details of what we see. The high-level conclusions are best moved to the main text} a) Dataset 5: Mouse - Phosphorylation - 2mix-TMT. Comparing the global profiling of protein $TTP$ with the modification at site $S178$. The modification and unmodified protein abundance followed the same trend across conditions. Specifically, there was a positive adjustment in abundance when comparing WT\_Uninfect to WT\_Late in both the modification and global profiling run. b) Dataset 4: Human - Ubiquitination - 1mix-TMT. Comparing the global profiling of protein $GSDMD$ with the ubiquitination of the protein at site $K62$. Here the modification and global protein follow different trends. There appeared to be no change in abundance between Dox1hr and Dox4hr in the modified plot, however there was a large negative change when looking at the unmodified plot. This indicated the modification was confounded with changes in the unmodified protein.}
\label{fig:profile_plots}
\end{figure}

\begin{figure}[h!]
\centering
 \begin{subfigure}{\textwidth}
 \centering
	\includegraphics[height=.55\textwidth]{images/usp30_venn_diagramm}
	\caption{}
	\label{fig:data6_vd1}
 \end{subfigure}
 \begin{subfigure}{\textwidth}
 \centering
	\includegraphics[height=.55\textwidth]{images/usp30_venn_diagramm_matching_only}
	\caption{}
	\label{fig:data6_vd2}
 \end{subfigure}
 \caption{Dataset 6: Human - Ubiquitination - Label-free no global profiling run. a) The overlap of differential modified peptides for the PTM model with and without global protein level adjustment across all comparisons. More PTMs became insignificant than became significant after adjustment. In total, 2709 modified peptides became insignificant, while only 547 became significant. b) Here we made the same comparison but only looked at modified peptides where adjustment could be performed, ie they had a matching unmodified protein. In this case there were significantly less peptides that became insignificant after adjustment. 726 modified peptides became insignificant, 547 became significant, and 1,078 were significant in both models.}
\label{fig:data6_plots}
\end{figure}

>>>>

$MSstatsPTM$ before adjusting for changes in the unmodified protein corresponds to base $MSstats$ or $MSstatsTMT$, modeled on the peptide level as described in {\bf Supplementary Table \ref{supp-fig:statistical_inference}}, as appropriate for the experimental design. $MSstatsPTM$ with the adjustment, described in \secref{adjustment_summary}, was used without imputing missing values, and without Empirical Bayes moderation.

$Unadjusted$ $ANOVA$, i.e. $ANOVA$ before adjusting for changes in the unmodified protein, was as in \eqref{ttest_model}. $Adjusted$ $ANOVA$, i.e. $ANOVA$ with the adjustment was modeled as in \eqref{ttest_ratio_model}. 
Finally, $unadjusted$ $Limma$ used the same model formula in \eqref{ttest_model}, while including a moderated variance estimation. $Adjusted$ $Limma$ was modeled as in \eqref{ttest_ratio_model} including moderated variance estimation.  
All the evaluations were done at the FDR-adjusted p-value cutoff of $q=.05$. More details are in {\bf Supplementary Sec. \ref{supp-sec:experiments}}.

@@

\figref{sim1_fdr} summarizes the eFDR reported on Computer Simulation 1 dataset by $MSstatsPTM$, $adjusted$ $ANOVA$, $adjusted$ $Limma$ and base $MSstats$, $unadjusted$ $ANOVA$ and $unadjusted$ $Limma$ methods. The simulation mimicked a ``clean" label-free group comparison experiment, not compromised by issues such as deviations from model assumptions, missing values and outliers. All the analyses were performed to control the eFDR at at most 5\%.  Yet, even under these favorable circumstances, the models that did not adjust for confounding from changes in overall protein abundance produced an excessive number of false positives. The versions of the models that accounted for the confounding produced error rates that were much better calibrated at the desired level.

%Since the simulation mimicked the assumptions of model XXX?, the error rates of this model were calibrated best.

@@

\figref{sim2_acc} summarizes the overall accuracy of methods adjusting for changes in the unmodified protein for Computer Simulation 2. The simulation mimicked a more realistic label-free group comparison experiment, including low counts of modified features and missing values. We evaluated the impact of low versus high noise, of the number of biological replicates, and of the number of conditions. For low noise, $MSstatsPTM$ outperformed the existing methods across all conditions and number of replicates, with near 100\% accuracy when the replicates were high. As the noise increased the accuracy of all the methods decreased, however $MStatsPTM$ still outperformed the existing methods. The difference is primarily due to the robust nature of summarization in $MSstatsPTM$ through Tukey's Median Polish. 
The ratio-based summarization approach used by $Limma$ and $ANOVA$ requires measurements for both the PTM and unmodified protein. $MSstatsPTM$ in contrast can still leverage the information in the PTM or unmodified protein if one is missing. On average over all simulations, 3.94\% of the total unmodified protein data used by $MSstatsPTM$ was thrown out by $Limma$ and $ANOVA$ due to missing PTM data. No PTM data was thrown out due to missing protein data.

@@

\figref{spikein_boxplot} summarizes the evaluation on the label-free spike-in group comparison experiment, for all the methods, with and without adjusting for changes in the unmodified protein. 
Without adjusting for changes in unmodified protein abundance, all the approaches incorrectly estimated the log$_2$-fold change of the modified spike-in peptides.
After adjustment, the estimation was generally in line with the ground truth for all methods, however $MSstatsPTM$'s distribution of estimated fold changes was tighter. 
On average over all comparisons the inter-quartile range of the estimates by $MSstatsPTM$ was 32.86\% smaller than that by $Limma$ and $ANOVA$. 
As seen in the previous section, the robust summarization by $MSstatsPTM$ was particularly useful when the number of features was low. 
This was also the case in this experiment, with an average of $1.37$ features per PTM. Additionally, the ratio based summarization used by $Limma$ and $ANOVA$ discarded more data than in the simulations. The ratio based summarization lost 9.63\% of the unmodified protein and 1.94\% of the PTM data used by $MSstatsPTM$. This additional information lead $MSstatsPTM$ to a more accurate fold change estimation and better calibrated variance.

\figref{spikein_prop_volcano} details the detection of differentially abundant PTM for $MSstatsPTM$ and $Limma$, with and without the adjustment for changes in the unmodified protein, for the $mix3$-$mix2$ and $mix4$-$mix2$ comparisons.
As above, the log$_2$-fold changes of the spike-in peptides were only correctly estimated when accounting for changes in the unmodified protein abundance. 
Additionally, the background peptides, serving as the null model, show many false positives before adjustment. 
After adjustment the extent of false positives substantially decreased. 
Specifically, for $MSstatsPTM$ the number of false positives went from 20.88\% to 1.84\% after the adjustment, and for $Limma$ went from 26.04\% to 1.18\%.
While the proposed method and $Limma$ both correctly estimated the fold change of the spike-in peptides, using $Limma$ resulted in many large adjusted p-values, and lower sensitivity. 
This was mainly due to $Limma$ estimating a higher variance, even when including variance moderation.
On average, over all the PTMs in this experiment, the variance components estimated by $Limma$ were 35.7\% larger than for $MSstatsPTM$. Volcano plots for all methods and comparisons can be seen in {\bf Supplementary Section~\ref{supp-sec:benchmark}}.

@@

The results of Dataset 4: Human - Ubiquitination - 1mix-TMT are summarized in~\figref{data4_venn_diagram} and~\figref{data4_profile_plot}. \figref{data4_venn_diagram} shows the number of differentially abundant modified peptides before and after adjustment. Adjusting for changes in the unmodified protein caused more PTMs to change from differentially abundant to non-differentially abundant than vice versa. A question that must be addressed is if the decrease in differentially abundant peptides is due to the increased standard error that comes from adjustment, as in \eqref{prop_se}. This was tested by looking for modified peptides whose adjusted log fold change was within 10\% of the unadjusted log fold change but became statistically insignificant after adjustment. In other words, the fold change was the same between models but the standard error increased. When the test was applied to Dataset 4, only one PTM became non-differentially abundant due to an increase in standard error. Thus we can conclude that the drop off in differentially abundant PTMs was due to changes in global protein abundance, not simply increased model variance.

\figref{data4_venn_diagram} demonstrates that the unmodified protein was more likely to cause the observed changes in PTM abundance, rather than mask changes in the PTM. Conversely, cases where the unmodified protein masks changes in the PTM are more infrequent. In this case, the modification and unmodified protein exhibit contradictory changes in abundance, moving in opposite directions. One such case can be seen in~\figref{data4_profile_plot}. Luchetti et al.~\cite{LUCHETTI2021} showed that $GSDMD$ was actively degraded when IpaH7.8 expression was induced by dox treatment. Our reanalysis confirmed this, with the $GSDMD$ protein being down-regulated when Dox treatments reached the 4 and 6 hour marks. Conversely, ubiquitination of $GSDMD$ at site $K62$ up-regulated abundance between the same conditions. This up-regulation was originally confounded by the down-regulation of unmodified $GSDMD$, and made the modification appear to have little change between no Dox and Dox 4 and 6 hour conditions. The proposed approach accounted for this confounding and the modification was shown to be differentially abundant, with an absolute fold change of $1.415$ between the Dox 1 hour and Dox 4 hour conditions. This was an interesting result, as the modification contradicts what previous research showed for the unmodified protein. This effect can be thought of as a false negative and would have been challenging to observe without the proposed approach. There are many non-differentially abundant PTMs, and having to comb through each one individually to determine false negatives is generally an unfeasible task. The proposed approach removes the confounding on an experimental level, revealing modifications of interest that are entirely masked due to changes in the unmodified protein.

The results of Dataset 5: Mouse - Phosphorylation - 2mix-TMT are summarized in~\figref{data5_venn_diagram} and~\figref{data5_profile_plot}. \figref{data5_venn_diagram} showed that adjusting for changes in the unmodified protein caused more PTMs to change from differentially abundant to non-differentially abundant. Again we tested if this change was due to fold change adjustment or an increased standard error. When applied to Dataset 5, 548 PTMs became non-differentially abundant due to an increase in standard error (only 3.4\% of all PTMs that became non-differentially abundant).

\figref{data5_profile_plot} shows a case where the unmodified protein caused the observed changes in PTM abundance. This occurs when the protein exhibits a change in abundance while the modification does not. Originally the modification at $S178$ of protein $TTP$ was shown to be differentially abundant between KO\_Uninfect and KO\_Early, with a log fold change of $2.9$. However the unmodified protein was shown to contribute $2.014$ of this change, meaning that the modification only accounted for $.886$ after adjustment (seen in {\bf Supplementary Figure~\ref{supp-fig:No_Diff_Shigella_PTM}}). This caused the modification to go from differentially abundant to non-differentially abundant. These cases occur frequently and can be thought of as false positives. The proposed approach corrects for this confounding, allowing us to see the true fold change of PTMs, while excluding those that do not exhibit differential changes.

%Adjusting for changes in the unmodified protein caused more PTMs to change from differentially abundant to non-differentially abundant than vice versa. A question that must be addressed is if the decrease in differentially abundant peptides is due to the increased variance \todo{variance of what? estimation of the variance parameter? Clarify} that comes from adjustment. This was tested by looking for modified peptides whose adjusted log fold change was within 10\% of the unadjusted log fold change but became statistically insignificant after adjustment. In other words, the fold change was the same between models but variance increased. When the test was applied to Dataset 4, only one PTM became non-differentially abundant due to an increase in variance. When applied to Dataset 5, 548 PTMs became non-differentially abundant due to an increase in variance (only 3.4\% of all PTMs that became non-differentially abundant). Thus we can conclude that the drop off in differentially abundant PTMs was due to changes in global protein abundance, not simply increased model variance.

%\figref{venn_diagrams} demonstrates that the unmodified protein was more likely to cause the observed changes in PTM abundance, rather than mask changes in the PTM. The later case can occur when the unmodified protein exhibits a change in abundance, while the modification does not. Additionally, this case may come about due to the high variability of PTMs compared to the global protein. An example of this can be seen in \figref{data5_profile_plot}. Originally the modification at $S178$ of protein $TTP$ was shown to be differentially abundant between KO\_Uninfect and KO\_Early, with a log fold change of $2.9$. However the unmodified protein was shown to contribute $2.014$ of this change, meaning that the modification only accounted for $.886$ after adjustment (seen in {\bf Supplementary Figure~\ref{supp-fig:No_Diff_Shigella_PTM}}). This caused the modification to go from differentially abundant to non-differentially abundant. These cases occur frequently and can be thought of as false positives. The proposed approach corrects for this confounding, allowing us to see the true fold change of PTMs, while excluding those that do not exhibit differential changes.

@@

The number of differentially abundant PTMs before and after adjustment are shown in~\figref{data6_plots}. After adjusting for changes in the unmodified protein, there were fewer differentially modified peptides. However, this was mainly due to not having a global profiling run. When looking only at PTMs with a corresponding unmodified protein, the number of differentially modified PTMs exhibited a small drop, but was more consistent. As above, we tested if this drop in differentially abundant PTMs was due to an increase in standard error. Here only 25 PTMs became insignificant due to increased standard error. Without a global profiling run, the method was less effective at removing confounding with the unmodified protein. This was due to not being able to perform adjustment without a corresponding global protein. In this case adjustment caused us to toss out a large number of the PTMs in the original model. In cases without a global profiling run, the proposed method can still be used, but it should be used in conjunction with the unadjusted model so that information is not lost.

@@

\begin{figure}[ht]
\centering
\begin{subfigure}[c]{0.825\linewidth}
\includegraphics[width=1\textwidth]{images/spike_in_fc.png}
\caption{}
\label{fig:spikein_boxplot}
\end{subfigure}
\begin{subfigure}[c]{0.825\linewidth}
\includegraphics[width=1\textwidth]{images/spike_in_volcano.png}
\caption{}
\label{fig:spikein_prop_volcano}
\end{subfigure}
\caption{Dataset 3: Spike-in benchmark - Ubiquitination - Label-free. a) The distribution of spike-in peptides log$_2$ fold change estimated by $MSstatsPTM$, $Limma$, and $ANOVA$ with and without adjustment. The expected log$_2$ fold change is highlighted by a red `X'. Protein adjustment removed systematic differences from the expected log$_2$ fold change in all models. b) The statistical results of $MSstatsPTM$ and $Limma$ modeling the $mix3$-$mix2$ and $mix4$-$mix2$ comparisons before and after adjustment. The spike-in peptides are colored red, while the background peptides are highlighted grey. The solid horizontal line shows the adjusted pvalue cutoff of .05. The solid vertical line shows log$_2$ fold change of 0. The dashed vertical line shows the expected log$_2$ fold change of the spike-in peptides. Before adjustment the spike-in peptides (colored red) did not follow the expected log fold change; after adjustment, the spike-in peptides were more in line with expectation.}
\label{fig:spikein_volcano}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{\textwidth}
 \centering
\includegraphics[width=1.0\textwidth]{images/IpaH_prof_plot.png}
\caption{}
\label{fig:data4_profile_plot}
 \end{subfigure}
 \begin{subfigure}{\textwidth}
 \centering
	\includegraphics[width=1.0\textwidth]{images/No_Difference_Shigella_Profile_Plot}
	\caption{}
	\label{fig:data5_profile_plot}
	 \end{subfigure}
\caption{a) Dataset 4: Human - Ubiquitination - 1mix-TMT. Comparing the $GSDMD$ protein and its ubiquitination at site $K62$ across all conditions. The individual PTM and protein features are shown in grey, while the summarization is highlighted in red. b) Dataset 5: Mouse - Phosphorylation - 2mix-TMT. Comparing the profiling of protein $TTP$ with the phosphorylation at site $S178$. The individual PTM and protein features are shown in grey, while the summarization is highlighted in red. The plots are separated into TMT Mixtures 1 and 2.}
\label{fig:profile_plots}
\end{figure}

\begin{figure}[h!]
\centering
 \begin{subfigure}{\textwidth}
 \centering
	\includegraphics[height=.55\textwidth]{images/usp30_venn_diagramm}
	\caption{}
	\label{fig:data6_vd1}
 \end{subfigure}
 \begin{subfigure}{\textwidth}
 \centering
	\includegraphics[height=.55\textwidth]{images/usp30_venn_diagramm_matching_only}
	\caption{}
	\label{fig:data6_vd2}
 \end{subfigure}
 \caption{Dataset 6: Human - Ubiquitination - Label-free no global profiling run. a) The overlap of differential modified peptides for the PTM model with and without global protein level adjustment across all comparisons, including all measured PTMs. b) The overlap of differential modified peptides with and without global protein level adjustment, including only PTMs with a corresponding unmodified protein.}
\label{fig:data6_plots}
\end{figure}
